---
title: "Abalone_Week1"
author: "Raghu"
date: "September 7, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
To clean the Environment
```{r}
rm(list = ls(all=TRUE))
```
TO make sure that the files we read exist in the workig directory or not
```{r}
setwd("C:/Users/RaghuTeja/Desktop/intern/b29-interns-data-RaghutejaDhulipala-master/submissions/Week1")
```
TO read the data 
```{r}
mydata <- read.table("abalone.txt", header=FALSE, sep=",")
```
TO know the first 6 rows of the dataset
```{r}
head(mydata)
```
To know the Last 6 rows in the data
```{r}
tail(mydata)
```
To know the structure of the data.
```{r}
str(mydata)
```
To check for any NA values in the entire dataset
```{r}
sum(is.na(mydata))
```
To find the detailed information of each any every attribute.
```{r}
summary(mydata)
```
TO check the Standard Deviation in each attribute
```{r}
sd(mydata$V2)
sd(mydata$V3)
sd(mydata$V4)
sd(mydata$V5)
sd(mydata$V6)
sd(mydata$V7)
sd(mydata$V8)
sd(mydata$V9)

```
To know the no: of unique values in each attribute
```{r}
unique(mydata$V1)
unique(mydata$v2)
unique(mydata$V3)
unique(mydata$V4)
unique(mydata$V5)
unique(mydata$V6)
unique(mydata$V7)
unique(mydata$V8)
unique(mydata$V9)
```


```{r}
target<-mydata$V1
mydata$V1<-NULL
```

To get the correlation matrix for the numerical attributes
```{r}
res <- cor(mydata)
round(res, 2)
```
TO see the correlation between the attributes graphically
```{r}

library(corrplot)

corrplot(cor(mydata), method = "number")
```

```{r}
mat <- cor(mydata)
round(mat, 2)
```

mydata$V3=NULL
mydata$V6=NULL
mydata$V7=NULL

```{r}
mydata = cbind(mydata,target)
```
