---
title: "Untitled"
author: "Raghu"
date: "November 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To clean the Environment
```{r}
rm(list = ls(all=TRUE))
```

TO make sure that the files we read exist in the workig directory or not
```{r}
setwd("C:/Users/RaghuTeja/Desktop/intern/")
```
TO read the data 
```{r}
mydata <- read.table("abalone.txt", header=FALSE, sep=",")
mydata<-mydata[-2052,]
```
Adding 1.5 to the entire coloumn(V9) as mentioned in description, for age of the abalone
```{r}
mydata$V9<-mydata$V9+1.5
```
Using library dummies as we have one attribute with 3 classes.
```{r}
#install.packages("dummies")
library(dummies)

```
Dummying the cloumn V1
```{r}
dummy1<-dummy(mydata$V1)
```
Dropping the column V1 as we have already dummied the column
```{r}
mydata$V1<-NULL
```
to check the class of the dummied variable
```{r}
class(dummy1)
```
As the class of the dummy is in matrix we change to dataframe
```{r}
dummy1<-as.data.frame(dummy1)
```
Adding the dummy columns to mydata
```{r}
mydata<-cbind(dummy1,mydata)
```

Target V9 is stored in the target variable for standardizing the entire data
```{r}
target<-mydata$V9
mydata$V9<-NULL
```
Standardizing the data using scale function
```{r}
dim(mydata)
mydata<-scale(mydata[,1:10], center = TRUE, scale = TRUE)
class(mydata)#mydata is in matrix format
mydata<-as.data.frame(mydata)#so changing it into dataframe
mydata<-cbind(mydata,target)#adding the target back to the data
str(mydata)
```
Diving the data into test and train so as to run the model and predict the best model
```{r}
set.seed(100)

# the "sample()" function helps us to randomly sample 70% of the row indices of the dataset

train_rows <- sample(x = 1:nrow(mydata), size = 0.7*nrow(mydata))

# We use the above indices to subset the train and test sets from the data

train_data <- mydata[train_rows, ]

test_data <- mydata[-train_rows, ]

```
Runing the linear model on the entire data initially. 
```{r}
model1 <- lm(formula = target ~., data = train_data)
summary(model1)
```
Plotting the linear model on the train data.
```{r}
par(mfrow = c(2,2)) 
plot(model1)
```
on train data
```{r}
preds <- predict(model1, train_data[, !(names(test_data) %in% c("target"))])

```
```{r}
mae(train_data[, "target"], preds)

mse(train_data[, "target"], preds)

rmse(train_data[, "target"], preds)

mape(train_data[, "target"], preds)

```

Predicting the model on the test data
```{r}
preds <- predict(model1, test_data[, !(names(test_data) %in% c("target"))])
```

```{r}
mae <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(abs(error))
  
}

mse <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(error^2)
  
}
rmse <- function(actual, predicted){
  
  error <- actual - predicted
  
  sqrt(mean(error^2))
  
}
mape <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(abs(error/actual))*100
  
}

mae(test_data[, "target"], preds)

mse(test_data[, "target"], preds)

rmse(test_data[, "target"], preds)

mape(test_data[, "target"], preds)
```
#KNN
```{r}
library(class)
library(DMwR)
library(caret)
model_knn <- knn3(target~.,train_data,k=3)
pred_knn <- predict(model_knn, train_data[, !(names(train_data) %in% c("target"))])
mae(train_data[, "target"], pred_knn)

mse(train_data[, "target"], pred_knn)

rmse(train_data[, "target"], pred_knn)

#On test data
pred_knn <- predict(model_knn, test_data[, !(names(test_data) %in% c("Target"))])
  
mae(test_data[, "target"], pred_knn)

mse(test_data[, "target"], pred_knn)

rmse(test_data[, "target"], pred_knn)

regr.eval(test_data$target, pred_knn)

```
CART
```{r}
library(rpart)

model_dt <- rpart(target ~ . , train_data,cp=0.01)
#On test data

pred_dt <- predict(model_dt, train_data[, !(names(train_data) %in% c("target"))])

mae(train_data[, "target"], pred_dt)

mse(train_data[, "target"], pred_dt)

rmse(train_data[, "target"], pred_dt)

#On test data
pred_dt <- predict(model_dt, test_data[, !(names(test_data) %in% c("target"))])

mae(test_data[, "target"], pred_dt)

mse(test_data[, "target"], pred_dt)

rmse(test_data[, "target"], pred_dt)

regr.eval(test_data$Target, pred_dt)


```
#SVM
```{r}
library(e1071)
library(caret)
model_svm <- svm(target ~ ., data = train_data, kernal = "linear",cost=1)
summary(model_svm)
#On train data
preds_svm <- predict(model_svm, train_data[, !(names(train_data) %in% c("target"))])

mae(train_data[, "target"], preds_svm)

mse(train_data[, "target"], preds_svm)

rmse(train_data[, "target"], preds_svm)
#On test data
preds_svm <- predict(model_svm, test_data[, !(names(test_data) %in% c("target"))])

mae(test_data[, "target"], preds_svm)

mse(test_data[, "target"], preds_svm)

rmse(test_data[, "target"], preds_svm)

regr.eval(test_data$Target, preds_svm)
```
#Random forest

```{r}
library(randomForest)
rf <- randomForest(target ~ ., data=train_data)
importance(rf)
#On train data
preds_rf_imp <- predict(rf, train_data[, !(names(train_data) %in% c("Target"))])
  
mae(train_data[, "target"], preds_rf_imp)

mse(train_data[, "target"], preds_rf_imp)

rmse(train_data[, "target"], preds_rf_imp)

#On test data
preds_rf_imp <- predict(rf, test_data[, !(names(test_data) %in% c("Target"))])
  
mae(test_data[, "target"], preds_rf_imp)

mse(test_data[, "target"], preds_rf_imp)

rmse(test_data[, "target"], preds_rf_imp)

regr.eval(test_data$target, preds_rf_imp)

```
