---
title: "Untitled"
author: "Raghu"
date: "October 16, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To clean the Environment
```{r}
rm(list = ls(all=TRUE))
```

TO make sure that the files we read exist in the workig directory or not
```{r}
setwd("C:/Users/RaghuTeja/Desktop/intern/")
```
TO read the data 
```{r}
mydata <- read.table("abalone.txt", header=FALSE, sep=",")
#mydata<-mydata[-2052,]
```
Adding 1.5 to the entire coloumn(V9) as mentioned in description, for age of the abalone
```{r}
mydata$V9<-mydata$V9+1.5
```
Using library dummies as we have one attribute with 3 classes.
```{r}
#install.packages("dummies")
library(dummies)

```
Dummying the cloumn V1
```{r}
dummy1<-dummy(mydata$V1)
```
Dropping the column V1 as we have already dummied the column
```{r}
mydata$V1<-NULL
```
to check the class of the dummied variable
```{r}
class(dummy1)
```
As the class of the dummy is in matrix we change to dataframe
```{r}
dummy1<-as.data.frame(dummy1)
```
Adding the dummy columns to mydata
```{r}
mydata<-cbind(dummy1,mydata)
```

Target V9 is stored in the target variable for standardizing the entire data
```{r}
target<-mydata$V9
mydata$V9<-NULL
```
Standardizing the data using scale function
```{r}
dim(mydata)
mydata<-scale(mydata[,1:10], center = TRUE, scale = TRUE)
class(mydata)#mydata is in matrix format
mydata<-as.data.frame(mydata)#so changing it into dataframe
mydata<-cbind(mydata,target)#adding the target back to the data
str(mydata)
```
Diving the data into test and train so as to run the model and predict the best model
```{r}
set.seed(100)

# the "sample()" function helps us to randomly sample 70% of the row indices of the dataset

train_rows <- sample(x = 1:nrow(mydata), size = 0.7*nrow(mydata))

# We use the above indices to subset the train and test sets from the data

train_data <- mydata[train_rows, ]

test_data <- mydata[-train_rows, ]

```
Runing the linear model on the entire data initially. 
```{r}
model1 <- lm(formula = target ~., data = train_data)
summary(model1)
```
Plotting the linear model on the train data.
```{r}
par(mfrow = c(2,2)) 
plot(model1)
```

```{r}
preds <- predict(model1, train_data[, !(names(test_data) %in% c("target"))])

```


Predicting the model on the test data
```{r}
preds <- predict(model1, test_data[, !(names(test_data) %in% c("target"))])
```

```{r}
mae <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(abs(error))
  
}

mse <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(error^2)
  
}
rmse <- function(actual, predicted){
  
  error <- actual - predicted
  
  sqrt(mean(error^2))
  
}
mape <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(abs(error/actual))*100
  
}

mae(test_data[, "target"], preds)

mse(test_data[, "target"], preds)

rmse(test_data[, "target"], preds)

mape(test_data[, "target"], preds)
```

#Step AIC

```{r}
library(MASS)
model_aic <- stepAIC(model1, direction = "both")

summary(model_aic)

par(mfrow = c(2,2))

plot(model_aic)

```
#VIF
```{r}
#install.packages("car")
library(car)

#vif(model1)

vif(model_aic)

```
By the VIF values checking the correlation between V5 &V6
```{r}
library(corrplot)
corrplot(cor(mydata))
cor(mydata$V5,mydata$V6)
mydata$V5<-NULL
train_data$V5<-NULL
test_data$V5<-NULL
str(train_data)
```

```{r}
#Without V5
model2 <- lm(formula = target ~., data = train_data)
summary(model2)

```

Predicting the model on the test data
```{r}
preds <- predict(model2, test_data[, !(names(test_data) %in% c("target"))])
```

```{r}
mae <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(abs(error))
  
}

mse <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(error^2)
  
}
rmse <- function(actual, predicted){
  
  error <- actual - predicted
  
  sqrt(mean(error^2))
  
}
mape <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(abs(error/actual))*100
  
}

mae(test_data[, "target"], preds)

mse(test_data[, "target"], preds)

rmse(test_data[, "target"], preds)

mape(test_data[, "target"], preds)
```
removing V6- Second highest VIF
```{r}
mydata$V6<-NULL
train_data$V6<-NULL
test_data$V6<-NULL

```

```{r}
#Without V5 and V6
model3 <- lm(formula = target ~., data = train_data)
summary(model3)

```

Predicting the model on the test data
```{r}
preds <- predict(model3, test_data[, !(names(test_data) %in% c("target"))])
```
```{r}
mae <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(abs(error))
  
}

mse <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(error^2)
  
}
rmse <- function(actual, predicted){
  
  error <- actual - predicted
  
  sqrt(mean(error^2))
  
}
mape <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(abs(error/actual))*100
  
}

mae(test_data[, "target"], preds)

mse(test_data[, "target"], preds)

rmse(test_data[, "target"], preds)

mape(test_data[, "target"], preds)
```

removing V7- third highest VIF
```{r}
mydata$V7<-NULL
train_data$V7<-NULL
test_data$V7<-NULL

```

```{r}
#Without V5, v6 and V7
model4 <- lm(formula = target ~., data = train_data)
summary(model4)

```

Predicting the model on the test data
```{r}
preds <- predict(model4, test_data[, !(names(test_data) %in% c("target"))])
```

```{r}
mae <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(abs(error))
  
}

mse <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(error^2)
  
}
rmse <- function(actual, predicted){
  
  error <- actual - predicted
  
  sqrt(mean(error^2))
  
}
mape <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(abs(error/actual))*100
  
}

mae(test_data[, "target"], preds)

mse(test_data[, "target"], preds)

rmse(test_data[, "target"], preds)

mape(test_data[, "target"], preds)
```
removing V8- fourth highest VIF
```{r}
mydata$V8<-NULL
train_data$V8<-NULL
test_data$V8<-NULL

```

```{r}
#Without V5, v6, v7 and V8
model5 <- lm(formula = target ~., data = train_data)
summary(model5)

```

Predicting the model on the test data
```{r}
preds <- predict(model5, test_data[, !(names(test_data) %in% c("target"))])
```

```{r}
mae <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(abs(error))
  
}

mse <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(error^2)
  
}
rmse <- function(actual, predicted){
  
  error <- actual - predicted
  
  sqrt(mean(error^2))
  
}
mape <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(abs(error/actual))*100
  
}

mae(test_data[, "target"], preds)

mse(test_data[, "target"], preds)

rmse(test_data[, "target"], preds)

mape(test_data[, "target"], preds)
```

```{r}

```

